### 2018-08-24 augsburg/talks/trento18/euroscipy.tex
first parts of EuroScipy talk

% pdflatex presentation

\documentclass[svgnames]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[sfdefault,scaled=.85,lining]{FiraSans}
\usepackage[scaled=0.85,lining]{FiraMono}
\usepackage{newtxsf}
\usepackage{rotating}
\usepackage{listings}
\usepackage{hyperref}

\usetheme{default}
\setbeamertemplate{navigation symbols}
{%
%  \hspace{3em}
%  \vbox{%
%  \hbox{\insertslidenavigationsymbol}
%  \hbox{\insertframenavigationsymbol}
%  \hbox{\insertbackfindforwardnavigationsymbol}
%  \vspace{2em}}
}

\setbeamerfont{frametitle}{family=\sffamily\firamedium}
\setbeamercolor{alerted text}{fg=red!70!black}
\setbeamercolor{structure}{fg=Navy}

\hypersetup{%
  pdftitle={Teaching programming with Jupyterhub and Nbgrader}
  ,pdfauthor={Gert-Ludwig Ingold <gert.ingold@physik.uni-augsburg.de>}
  ,pdfsubject={Talk at EuroSciPy 2018, Trento 31.8.2018}
  ,pdfkeywords={Python, Jupyterhub, nbgrader, education, EuroSciPy}
}

\definecolor{positive}{rgb}{0, 0.5, 0}
\definecolor{negative}{rgb}{0.7, 0, 0}
\definecolor{myred}{rgb}{0.8, 0, 0}
\definecolor{mygreen}{rgb}{0, 0.6, 0}
\definecolor{myblue}{rgb}{0, 0, 0.8}

\graphicspath{{./images/}}

\begin{document}

\begin{frame}[t]
 \vspace{1.5truecm}
 \begin{center}
  \structure{\huge\textbf{Teaching programming with}}\\[0.2truecm]
  \structure{\huge\textbf{Jupyterhub and Nbgrader}}\\[0.8truecm]
  \structure{\Large Gert-Ludwig Ingold}\\[0.1truecm]
  \structure{\large Universität Augsburg}
 \end{center}
\end{frame}

\begin{frame}{About me}
 \begin{itemize}
  \item theoretical physicist at the Universität Augsburg
  \item teaching programming to physicists and materials scientists since 2010
  \item involved in two Erasmus+ projects on education and computing
	\begin{itemize}
         \item iCSE4school (2015--2017)
	 \item Juypter@edu (2017--2019)
	\end{itemize}
 \end{itemize}

 \vspace{0.8truecm}
 \structure{Jupyter@edu}\qquad \raisebox{-0.2truecm}{\includegraphics[width=0.25\textwidth]{EU_flag-Erasmus__vect_POS}}

 \begin{itemize}
  \item University of Silesia, Poland (Leader)
  \item Universität Augsburg, Germany
  \item Universidade Portucalense Infante D. Henrique, Portugal
  \item European University Cyprus, Cyprus
 \end{itemize}
\end{frame}

\begin{frame}{Grading}
 \includegraphics[width=\textwidth]{function}

 \includegraphics[width=\textwidth]{tests}
\end{frame}

\begin{frame}{Functions}
 the students' code needs to be structured in functions to be tested
 \begin{itemize}
  \item concept of a function needs to be introduced very early in the course\\
	but we will always have well defined inputs and outputs\\
	the more special aspects of a function (no argument, default arguments,
	keyword arguments, \dots) are not required
  \item students get used to logically structured code from the very beginning\\
	but they do not go through the routine of structuring the code by themselves
  \item docstrings are required to define what the function is supposed to do\\
	students get used early to the idea that a function contains a docstring\\
	but they are not writing the docstring by themselves
 \end{itemize}
\end{frame}

\begin{frame}{Grading with tests}
 \begin{itemize}
  \item Test-driven development\\
	but: tests are not developed by the students
  \item Tests allow to give feedback through error messages\\
	but: students tend to rely on this feedback instead of developing their own critical view on their code\\
	potential problem of incomplete tests
  \item ``Stupid mistakes'' will be made even when students get more advanced\\
	also test for the simple things: are results returned, do they have the correct type, \dots\\
	``trivial'' standard tests + tests of specific functionality
  \item The test should not disclose the solution.
 \end{itemize}
\end{frame}

\begin{frame}{Coping with heterogeneity}
 \begin{itemize}
  \item students without programming experience and (sometimes) code writing apprehension
  \item students with knowledge in another programming language\\
	code often does not look very pythonic, example: loop over objects vs. loop over indices
 \end{itemize}

 students with programming experience should not be bored and should obtain an interesting result from their code

 \begin{itemize}
  \item scientific applications are often considered as an extra mental burden
 \end{itemize}
\end{frame}

\begin{frame}{Examples of exercises}

 \vspace{-0.7truecm}
 \begin{columns}[t]
  \begin{column}{0.5\textwidth}
   \begin{center}
    \includegraphics[width=0.8\textwidth]{spiral}

    selected problems from\\[-0.05truecm] Project Euler (\url{projecteuler.net})
   \end{center}
  \end{column}%
  \begin{column}{0.5\textwidth}
   \begin{center}
    \includegraphics[width=0.8\textwidth]{julia}

    Julia set
   \end{center}
  \end{column}%
 \end{columns}

 \begin{columns}[t]
  \begin{column}{0.5\textwidth}
   \begin{center}
    \includegraphics[width=0.8\textwidth]{conway}

    Conway's game of life
   \end{center}
  \end{column}%
  \begin{column}{0.5\textwidth}
   \begin{center}
    \includegraphics[width=0.8\textwidth]{pi}

    $\pi$ to a few thousand digits
   \end{center}
  \end{column}%
 \end{columns}

 \vspace{0.4truecm}
 {\scriptsize\url{https://github.com/marcinofulus/jupyter4edu/tree/master/augsburg/exercises}}
\end{frame}

\begin{frame}{Branching paths}
 \only<1>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_1}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{JuliaSet_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<2>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_2}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{ColorRepresentation_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<3>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_3}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{Grid_screenshot}
   \end{column}
  \end{columns}
 }%
 \only<4>{%
  \begin{columns}
   \begin{column}{0.3\textwidth}
    \includegraphics[height=0.8\textheight]{branching_4}
   \end{column}%
   \begin{column}{0.7\textwidth}
    \includegraphics[width=\textwidth]{JuliaIteration_screenshot}
   \end{column}
  \end{columns}
 }
\end{frame}

\begin{frame}{Branching paths}
 \begin{itemize}
  \item individual path through problem possible
  \item other notebooks are opened in new tabs, it is easy to lose track
 \end{itemize}
\end{frame}

\end{document}

### 2018-08-24 augsburg/talks/trento18/images/src/branching.py

@@ -1,0 +1,64 @@
+import sys, os
+from pyx import canvas, color, deco, deformer, path, text, trafo, unit
+
+def set_myattrs(n, nhl):
+    normalcolor = color.hsb(0.66, 1, 0.6)
+    highlight = color.hsb(0, 1, 0.6)
+    if n == nhl:
+        return [highlight]
+    else:
+        return [normalcolor]
+
+def file(c, size=1, xoff=0, yoff=0, attrs=[], title=''):
+    w = 3
+    h = 2.1
+    fold = 0.6
+    outline = path.path(path.moveto(0, 0),
+                        path.lineto(w, 0),
+                        path.lineto(w, h-fold),
+                        path.lineto(w-fold, h),
+                        path.lineto(0, h),
+                        path.closepath())
+    foldpath = path.path(path.moveto(w-fold, h),
+                         path.lineto(w-fold, h-fold),
+                         path.lineto(w, h-fold))
+    c1 = canvas.canvas()
+    c1.stroke(outline, attrs)
+    d = deformer.smoothed(0.2)
+    c1.stroke(d.deform(foldpath), attrs)
+    c1.stroke(path.rect(0.1*w, 0.3*h, 0.6*w, 0.1*h))
+    c1.stroke(path.rect(0.1*w, 0.45*h, 0.6*w, 0.1*h))
+    c1.stroke(path.rect(0.1*w, 0.6*h, 0.6*w, 0.1*h))
+    c1.stroke(path.rect(0.1*w, 0.75*h, 0.6*w, 0.1*h))
+    c1.text(0.1, 0.1, r'\sffamily '+title, [trafo.scale(0.7)])
+    myattrs = [trafo.translate(xoff, yoff), trafo.scale(size)]
+    myattrs.extend(attrs)
+    c.insert(c1, myattrs)
+
+text.set(text.LatexRunner)
+text.preamble(r'''%\usepackage[sfdefault,scaled=0.85,lining]{FiraSans}
+                  \usepackage{arev}
+                  \usepackage[utf8]{inputenc}''')
+unit.set(wscale=1.3)
+
+name = os.path.splitext(sys.argv[0])[0]+'_{}'
+for norder in range(4):
+    c = canvas.canvas()
+    file(c, yoff=2.5, title='JuliaSet',
+         attrs=set_myattrs(norder, 0))
+    file(c, xoff=1, yoff=0, title=r'$\sim${}ColorRepresentation',
[…]

### 2018-08-24 augsburg/talks/trento18/images/src/conway.py

@@ -1,0 +1,59 @@
+from pyx import canvas, color, deformer, graph, path, trafo
+import numpy as np
+from scipy import signal
+
+class Conway:
+    """Conway's game of life
+
+    """
+    v = np.array([[1, 1, 1],
+                  [1, 0, 1],
+                  [1, 1, 1]])
+
+    def __init__(self, size=100):
+        """prepare initial state of game of life
+
+           size: Size of the square lattice on which the population
+                 dynamics unfolds. Periodic boundary conditions are
+                 used.
+
+        """
+        self.size = size
+        self.world = np.random.randint(0, 2, (self.size, self.size))
+
+    def update(self, *args):
+        """update the population in self.world
+
+        """
+        n = signal.convolve2d(self.world, self.v, mode='same', boundary='wrap')
+        self.world = self.world & (n == 2)
+        self.world = self.world | (n == 3)
+
+c = canvas.canvas()
+dx = 0.3
+h = 4
+w = 6.5
+p = path.rect(-dx, -dx, w+2*dx, h+2*dx)
+p = deformer.smoothed(0.5).deform(p)
+c.fill(p, [color.grey(0.5), trafo.translate(0.05, -0.05)])
+
+c1 = canvas.canvas([canvas.clip(p)])
+c1.fill(p, [color.grey(0.9)])
+
+np.random.seed = 42
+game = Conway(300)
+for _ in range(15):
+    game.update()
+world = game.world
+xoff = -0.5*w-dx
+yoff = -0.5*h-dx
+delta = 0.1
+for nx in range(game.size):
+    for ny in range(game.size):
+        if world[nx, ny]:
+            p = path.rect(xoff+(nx+0.1)*delta, yoff+(ny+0.1)*delta,
+                          0.8*delta, 0.8*delta)
+            c1.fill(p, [color.hsb(0.25, 1, 0.5)])
+
+c.insert(c1)
+c.writePDFfile()

### 2018-08-24 augsburg/talks/trento18/images/src/julia.py

@@ -1,0 +1,59 @@
+from pyx import canvas, color, deformer, graph, path, trafo
+import numpy as np
+
+def julia_iteration(x, y, c, schwelle, maxit):
+    """Führe die Julia-Iterationsvorschrift aus, bis der
+       Betrag des Ergebnisses den Wert von schwelle
+       überschreitet. Es werden maximal maxit Iterationen
+       ausgeführt.
+
+       Die Iterationsvorschrift lautet z→z²+c, wobei
+       z=x+iy.
+
+       Es wird die Zahl der benötigten Iterationen bis zum
+       Erreichend der Schwelle zurückgegeben. Wird diese
+       nicht erreicht, so wird maxit zurückgegeben.
+
+    """
+    z = complex(x, y)
+    for n in range(maxit):
+        z = z**2+julia_c
+        if abs(z) >= schwelle:
+            break
+    return n+1
+
+xmin = -0.3
+xmax = -0.1
+ymin = 0.4
+ymax = 0.6
+seitenlaenge = 1000
+julia_c = -0.8+0.156j
+schwelle = 50000
+maxit = 600
+
+julia_daten = []
+
+for y in np.linspace(ymin, ymax, seitenlaenge):
+    for x in np.linspace(xmin, xmax, seitenlaenge):
+        iterationen = julia_iteration(x, y, julia_c,
+                                      schwelle, maxit)
+        julia_daten.append((x, y, iterationen/maxit))
+
+c = canvas.canvas()
+dx = 0.3
+h = 4
+w = 6.5
+p = path.rect(-dx, -dx, w+2*dx, h+2*dx)
+p = deformer.smoothed(0.5).deform(p)
+c.fill(p, [color.grey(0.5), trafo.translate(0.05, -0.05)])
+
+c1 = canvas.canvas([canvas.clip(p)])
+g = graph.graphxy(height=8, width=8,
+                  x=graph.axis.linear(min=xmin, max=xmax),
+                  y=graph.axis.linear(min=ymin, max=ymax))
+g.plot(graph.data.points(julia_daten, x=1, y=2, color=3, title="iterations"),
+       [graph.style.density(gradient=color.rgbgradient.Rainbow)])
+c1.insert(g, [trafo.translate(-0.1*w, -0.5*h)])
+
+c.insert(c1)
+c.writePDFfile()

### 2018-08-24 augsburg/talks/trento18/images/src/pi.py

@@ -1,0 +1,26 @@
+from pyx import canvas, color, deformer, path, text, trafo, unit
+
+text.set(text.LatexRunner)
+text.preamble(r'''\usepackage[scaled=0.85,lining]{FiraMono}
+                  \usepackage[utf8]{inputenc}''')
+pistring = '3.14159265358979323846264338327950288419716939937510582097494459230781640628620899862803482534211706798214808651328230664709384460955058223172535940812848111745028410270193852'
+char_per_line = 25
+pistring = ' '.join([pistring[n*char_per_line:(n+1)*char_per_line]
+                     for n in range(10)])
+unit.set(xscale=1.45)
+
+c = canvas.canvas()
+dx = 0.3
+h = 4
+w = 6.5
+p = path.rect(-dx, -dx, w+2*dx, h+2*dx)
+p = deformer.smoothed(0.5).deform(p)
+c.fill(p, [color.grey(0.5), trafo.translate(0.05, -0.05)])
+c.fill(p, [color.grey(0.9)])
+c.text(0, 0, r'\noindent\texttt{{{}}}'.format(pistring),
+       [text.valign.bottom, text.parbox(5)])
+c.text(0.5*w, 0.5*h, r'\Huge $\pi$', [text.halign.center, text.valign.middle,
+                                      color.hsb(0.66, 0.8, 1), color.transparency(0.2),
+                                      trafo.scale(5.8)])
+
+c.writePDFfile()

### 2018-09-05 augsburg/talks/trento18/euroscipy.tex

@@
 \setbeamercolor{structure}{fg=Navy}
+\setbeamertemplate{section in toc shaded}[default][30]
 
@@
 \definecolor{myblue}{rgb}{0, 0, 0.8}
+
+\newcommand\but{\alert{\textit{but}} }
 
@@
   \item theoretical physicist at the Universität Augsburg
-  \item teaching programming to physicists and materials scientists since 2010
+  \item teaching programming to physicists and materials scientists\\
+	since 2010, typically 10+ students
+  \item in two years: mandatory course for 100+ students $\rightarrow$
+	  \structure{nbgrader}
   \item involved in two Erasmus+ projects on education and computing
@@
  \vspace{0.8truecm}
- \structure{Jupyter@edu}\qquad \raisebox{-0.2truecm}{\includegraphics[width=0.25\textwidth]{EU_flag-Erasmus__vect_POS}}
+ \structure{Jupyter@edu}\hspace{4.7truecm}%
+ \raisebox{-0.7truecm}{\includegraphics[width=0.37\textwidth]{EU_flag-Erasmus__vect_POS}}
 
+ \vspace{-0.7truecm}
  \begin{itemize}
@@
 
+\begin{frame}{nbgrader}
+ \begin{center}
+ \includegraphics[width=\textwidth]{nbgrader}
+ \end{center}
+
+ \url{nbgrader.readthedocs.io}\qquad \url{github.com/jupyter/nbgrader/}
+\end{frame}
+
+\begin{frame}
+ \begin{center}
+  \begin{minipage}{0.6\textwidth}
+   \tableofcontents
+  \end{minipage}
+ \end{center}
+\end{frame}
+
+\section{Jupyterhub and Jupyter notebooks}
+
+\begin{frame}
+ \begin{center}
+  \begin{minipage}{0.6\textwidth}
+   \tableofcontents[currentsection]
+  \end{minipage}
+ \end{center}
+\end{frame}
+
+\begin{frame}{Jupyterhub and nbgrader}
+
+ \begin{center}
+ \includegraphics[width=0.9\textwidth]{jupyterhub_nbgrader}
+ \end{center}
+ \begin{itemize}
+  \item easily accessible interface to problem sets
+  \item no need to install Python on local computer
+  \item consistent working environment for all students
+  \item \but may be beginners should gather experience with Python
+	on their own computer
+ \end{itemize}
+\end{frame}
+
+\begin{frame}{Jupyter notebooks or something else?}
+ \begin{columns}
+  \begin{column}{0.6\textwidth}
+   \includegraphics[width=\textwidth]{notebook}
+  \end{column}%
+  \begin{column}{0.4\textwidth}
+   \includegraphics[width=\textwidth]{spyder}
+
+   \includegraphics[width=\textwidth]{vim}
+  \end{column}
+ \end{columns}
+
+ \vspace{0.5truecm}
+ \begin{itemize}
+  \item notebook allows to guide students through a problem set
+  \item students might think that programming in Python implies
+	working with a Jupyter notebook
+ \end{itemize}
+\end{frame}
+
+\section{grading}
+
+\begin{frame}
+ \begin{center}
+  \begin{minipage}{0.6\textwidth}
+   \tableofcontents[currentsection]
+  \end{minipage}
+ \end{center}
+\end{frame}
+
 \begin{frame}{Grading}
- \includegraphics[width=\textwidth]{function}
-
- \includegraphics[width=\textwidth]{tests}
+ \begin{center}
+  \includegraphics[width=0.8\textwidth]{answer_tests}
+ \end{center}
+ \begin{itemize}
+  \item autograded answer
+  \item autograder tests
+  \item manually graded answer
+ \end{itemize}
 \end{frame}
@@
 \begin{frame}{Functions}
- the students' code needs to be structured in functions to be tested
+ \begin{center}
+  \includegraphics[width=0.7\textwidth]{answer_tests_function}
+ \end{center}
+
+ \vspace{-0.4truecm}
  \begin{itemize}
-  \item concept of a function needs to be introduced very early in the course\\
-	but we will always have well defined inputs and outputs\\
-	the more special aspects of a function (no argument, default arguments,
-	keyword arguments, \dots) are not required
-  \item students get used to logically structured code from the very beginning\\
-	but they do not go through the routine of structuring the code by themselves
-  \item docstrings are required to define what the function is supposed to do\\
-	students get used early to the idea that a function contains a docstring\\
-	but they are not writing the docstring by themselves
+  \item need to introduce functions very early in the course\\
+	\but special aspects (no arguments, no return value, default arguments,
+	keyword arguments, \dots) not needed
+  \item students get used to logically structured code early on\\
+	\but they do not do it themselves
+  \item include docstrings to make task well defined\\
+	students get used to the idea that a function contains a docstring\\
+        \but they are not writing the docstring by themselves
  \end{itemize}
@@
  \begin{itemize}
-  \item Test-driven development\\
-	but: tests are not developed by the students
-  \item Tests allow to give feedback through error messages\\
-	but: students tend to rely on this feedback instead of developing their own critical view on their code\\
-	potential problem of incomplete tests
-  \item ``Stupid mistakes'' will be made even when students get more advanced\\
-	also test for the simple things: are results returned, do they have the correct type, \dots\\
-	``trivial'' standard tests + tests of specific functionality
-  \item The test should not disclose the solution.
+  \item test-driven development\\
+	\but tests are not developed by the students
+  \item tests allow to give feedback through error messages\\
+	\but students tend to rely on this feedback instead of developing their
+	own critical view on their code
+  \item ``Stupid mistakes'' will be made not only at the beginning
+	\begin{itemize}
+         \item ``trivial'' standard tests (are results returned, do they have the correct type, \dots?)
+	 \item + tests of specific functionality which should not disclose the solution
 	\end{itemize}
+ \end{itemize}
+ \begin{center}
+  \includegraphics[width=\textwidth]{answer_tests_unittest1}
+ \end{center}
+\end{frame}
+
+\section{branching paths}
+
+\begin{frame}
+ \begin{center}
+  \begin{minipage}{0.6\textwidth}
+   \tableofcontents[currentsection]
+  \end{minipage}
+ \end{center}
 \end{frame}
@@
   \item students with knowledge in another programming language\\
-	code often does not look very pythonic, example: loop over objects vs. loop over indices
+	code often does not look very pythonic\\
+	example: loop over objects vs. loop over indices
  \end{itemize}
@@
  \begin{itemize}
-  \item scientific applications are often considered as an extra mental burden
+  \item \but scientific applications are often considered as an extra mental burden
  \end{itemize}
@@
 
-\begin{frame}{Examples of exercises}
+\begin{frame}{Examples of problem sets}
 
@@
   \end{columns}
+ }%
+ \only<5>{%
+  \begin{columns}[b]
+   \begin{column}{0.3\textwidth}
+    \includegraphics[height=0.8\textheight]{branching_4}
+   \end{column}%
+   \begin{column}{0.7\textwidth}
+    \begin{itemize}
+     \item individual path through problem possible
+     \item \but notebooks are opened in new tabs, it is easy to lose track
+    \end{itemize}
+
+    \vspace{1truecm}
+   \end{column}
+  \end{columns}
  }
@@
 
-\begin{frame}{Branching paths}
- \begin{itemize}
-  \item individual path through problem possible
-  \item other notebooks are opened in new tabs, it is easy to lose track
- \end{itemize}
-\end{frame}
-
 \end{document}

### 2019-09-18 augsburg/talks/katowice19/images/src/primetime.py
first version of Katowice talk

@@ -1,0 +1,68 @@
+from math import cos, radians, sin, sqrt
+from pyx import canvas, color, deformer, path, style, text, trafo, unit
+
+def prime_list(maximum):
+    if maximum < 2:
+        raise ValueError('maximum should be at least 2')
+    candidates = [False, False]+[True]*(maximum-1)
+    maxcandidate = int(sqrt(maximum)+0.5)
+    candidate = candidates.index(True)
+    while candidate <= maxcandidate:
+        for multiple in range(candidate**2, maximum+1, candidate):
+            candidates[multiple] = False
+        offset = candidate+1
+        candidate = candidates[offset:].index(True)+offset
+    primelist = [nr for nr, isprime in enumerate(candidates) if isprime]
+    return primelist
+
+def is_time(n):
+    hours, minutes = divmod(n, 100)
+    return 0 <= hours <= 23 and 0 <= minutes <= 59
+
+def formattime(t):
+    s = '{:04}'.format(t)
+    return r'\scriptsize\texttt{{{}:{}}}'.format(s[0:2], s[2:])
+
+text.set(text.LatexRunner)
+text.preamble(r'''\usepackage[scaled=0.85,lining]{FiraMono}
+                  \usepackage[utf8]{inputenc}''')
+unit.set(xscale=1.2)
+
+c = canvas.canvas()
+dx = 0.3
+h = 4
+w = 6.5
+p = path.rect(-dx, -dx, w+2*dx, h+2*dx)
+p = deformer.smoothed(0.5).deform(p)
+c.fill(p, [color.grey(0.5), trafo.translate(0.05, -0.05)])
+c.fill(p, [color.grey(0.9)])
+
+c2 = canvas.canvas([canvas.clip(p)])
+primetimes = [p for p in prime_list(2359) if is_time(p)]
+s = text.text(0, 0, formattime(primetimes[0]))
+textwidth, textheight = s.width, s.height
+nx = int(w/unit.tocm(textwidth))
+xoff = 0.5*(w-1.2*textwidth*nx+0.2*textwidth)
+for n, p in enumerate(primetimes):
+    s = formattime(p)
+    x = xoff+(n % nx)*1.2*textwidth
+    y = 1.03*h-(n//nx)*textheight*1.4
[…]

### 2019-09-18 augsburg/talks/katowice19/ingold.tex

--- 2018-09-05 augsburg/talks/trento18/euroscipy.tex
@@
 \hypersetup{%
-  pdftitle={Teaching programming with Jupyterhub and Nbgrader}
+  pdftitle={Teaching coding with nbgrader - lessons learned}
   ,pdfauthor={Gert-Ludwig Ingold <gert.ingold@physik.uni-augsburg.de>}
-  ,pdfsubject={Talk at EuroSciPy 2018, Trento 31.8.2018}
-  ,pdfkeywords={Python, Jupyterhub, nbgrader, education, EuroSciPy}
+  ,pdfsubject={Talk at Jupyter4edu conference, Katowice 23.9.2019}
+  ,pdfkeywords={Python, Jupyterhub, nbgrader, education}
 }
@@
  \begin{center}
-  \structure{\huge\textbf{Teaching programming with}}\\[0.2truecm]
-  \structure{\huge\textbf{Jupyterhub and Nbgrader}}\\[0.8truecm]
+  \structure{\huge\textbf{Teaching coding with nbgrader}}\\[0.2truecm]
+  \structure{\huge\textbf{-- lessons learned --}}\\[0.8truecm]
   \structure{\Large Gert-Ludwig Ingold}\\[0.1truecm]
@@
 
-\begin{frame}{About me}
+\begin{frame}{Teaching situation}
  \begin{itemize}
-  \item theoretical physicist at the Universität Augsburg
+  \item programming language: Python\\
+	{\small Dlaczego warto uczyć w języku python? {\scriptsize$\nearrow$} \url{youtu.be/CEmg3gdFlaE}}\\
+	{\small Why teach Python? {\scriptsize$\nearrow$} \url{youtu.be/X18jt2lVmJY}}
+        \begin{itemize}
+	 \item low entry barrier
+	 \item expressive code
+	 \item well suited for serious work
+	 \item freely available, ``batteries included´´
+         \item Anaconda distribution for coherent environment on
+	       current operating systems
+	\end{itemize}
   \item teaching programming to physicists and materials scientists\\
 	since 2010, typically 10+ students
-  \item in two years: mandatory course for 100+ students $\rightarrow$
-	  \structure{nbgrader}
-  \item involved in two Erasmus+ projects on education and computing
-	\begin{itemize}
-         \item iCSE4school (2015--2017)
-	 \item Juypter@edu (2017--2019)
-	\end{itemize}
+  \item next year: mandatory course for 100+ students
  \end{itemize}
 
- \vspace{0.8truecm}
- \structure{Jupyter@edu}\hspace{4.7truecm}%
- \raisebox{-0.7truecm}{\includegraphics[width=0.37\textwidth]{EU_flag-Erasmus__vect_POS}}
+\end{frame}
 
- \vspace{-0.7truecm}
+\begin{frame}{Coping with heterogeneity}
  \begin{itemize}
-  \item University of Silesia, Poland (Leader)
-  \item Universität Augsburg, Germany
-  \item Universidade Portucalense Infante D. Henrique, Portugal
-  \item European University Cyprus, Cyprus
+  \item students without programming experience and (sometimes) code writing apprehension
+  \item students with knowledge in another programming language\\
+	code often does not look very pythonic\\
+	example: loop over objects vs. loop over indices
+ \end{itemize}
+
+ students with programming experience should not be bored and should obtain an interesting
+ result from their code
+
+ \begin{itemize}
+  \item \but scientific applications are often considered as an extra mental burden
  \end{itemize}
@@
 \begin{frame}{nbgrader}
- \begin{center}
- \includegraphics[width=\textwidth]{nbgrader}
- \end{center}
+ \begin{itemize}
+  \item tool that facilitates creating and grading assignments in the Jupyter notebook
+  \item distributing and collecting assignments
+  \item automatic grading by means of unit tests
+  \item manual grading possible
+  \item feedback can be given to students
+ \end{itemize}
 
- \url{nbgrader.readthedocs.io}\qquad \url{github.com/jupyter/nbgrader/}
+ \begin{itemize}
+  \item freely available
+  \item open source
+  \item latest version: 0.6.0 released end of August 2019
+ \end{itemize}
 \end{frame}
 
-\begin{frame}
+\begin{frame}{nbgrader on Github}
  \begin{center}
-  \begin{minipage}{0.6\textwidth}
-   \tableofcontents
-  \end{minipage}
+  \url{github.com/jupyter/nbgrader/}
+
+  \vspace{0.3truecm}
+  \includegraphics[width=\textwidth]{nbgrader_github}
  \end{center}
@@
 
-\section{Jupyterhub and Jupyter notebooks}
+\begin{frame}{Github issues for nbgrader}
+ \begin{center}
+  report and discuss problems or new features
 
-\begin{frame}
- \begin{center}
-  \begin{minipage}{0.6\textwidth}
-   \tableofcontents[currentsection]
-  \end{minipage}
+  \vspace{0.3truecm}
+  \includegraphics[width=\textwidth]{nbgrader_githubissues}
  \end{center}
@@
 
-\begin{frame}{Jupyterhub and nbgrader}
+\begin{frame}{Pull requests for nbgrader}
+ \begin{center}
+  contribute to the code or the documentation
 
+  \vspace{0.3truecm}
+  \includegraphics[width=\textwidth]{nbgrader_githubprs}
+ \end{center}
+\end{frame}
+
+\begin{frame}{nbgrader documentation}
  \begin{center}
- \includegraphics[width=0.9\textwidth]{jupyterhub_nbgrader}
+  \url{nbgrader.readthedocs.io}
+
+  \vspace{0.3truecm}
+  \includegraphics[width=\textwidth]{nbgrader_rtd}
  \end{center}
+\end{frame}
+
+\begin{frame}{Jupyter notebook and Jupyterhub -- pros and cons}
+
+ \structure{Jupyter notebook}
+ \begin{itemize}
+  \item notebook allows to guide students through a problem set
+  \item problems related to out-of-order execution
+  \item students might think that programming in Python
+	necessarily implies working with a Jupyter notebook,
+        they should know about IDEs (spyder, \dots) and
+	editors (vim, \dots)
+ \end{itemize}
+
+ \structure{Jupyterhub}
  \begin{itemize}
@@
 
-\begin{frame}{Jupyter notebooks or something else?}
- \begin{columns}
-  \begin{column}{0.6\textwidth}
-   \includegraphics[width=\textwidth]{notebook}
-  \end{column}%
-  \begin{column}{0.4\textwidth}
-   \includegraphics[width=\textwidth]{spyder}
-
-   \includegraphics[width=\textwidth]{vim}
-  \end{column}
- \end{columns}
-
- \vspace{0.5truecm}
- \begin{itemize}
-  \item notebook allows to guide students through a problem set
-  \item students might think that programming in Python implies
-	working with a Jupyter notebook
- \end{itemize}
-\end{frame}
-
-\section{grading}
-
-\begin{frame}
- \begin{center}
-  \begin{minipage}{0.6\textwidth}
-   \tableofcontents[currentsection]
-  \end{minipage}
- \end{center}
-\end{frame}
-
-\begin{frame}{Grading}
- \begin{center}
-  \includegraphics[width=0.8\textwidth]{answer_tests}
- \end{center}
- \begin{itemize}
-  \item autograded answer
-  \item autograder tests
-  \item manually graded answer
- \end{itemize}
-\end{frame}
-
 \begin{frame}{Functions}
  \begin{center}
-  \includegraphics[width=0.7\textwidth]{answer_tests_function}
+  \includegraphics[width=\textwidth]{examplecell}
  \end{center}
 
- \vspace{-0.4truecm}
+ \vspace{0.3truecm}
  \begin{itemize}
@@
 \begin{frame}{Grading with tests}
+ \begin{center}
+  \includegraphics[width=\textwidth]{tests}
+ \end{center}
+
+ \vspace{0.3truecm}
  \begin{itemize}
@@
 
-\section{branching paths}
-
-\begin{frame}
- \begin{center}
-  \begin{minipage}{0.6\textwidth}
-   \tableofcontents[currentsection]
-  \end{minipage}
- \end{center}
-\end{frame}
-
-\begin{frame}{Coping with heterogeneity}
- \begin{itemize}
-  \item students without programming experience and (sometimes) code writing apprehension
-  \item students with knowledge in another programming language\\
-	code often does not look very pythonic\\
-	example: loop over objects vs. loop over indices
- \end{itemize}
-
- students with programming experience should not be bored and should obtain an interesting result from their code
-
- \begin{itemize}
-  \item \but scientific applications are often considered as an extra mental burden
- \end{itemize}
-\end{frame}
-
-\begin{frame}{Examples of problem sets}
+\begin{frame}{Examples of problem sets I}
 
@@
 
-    selected problems from\\[-0.05truecm] Project Euler (\url{projecteuler.net})
+    a problem from\\[-0.05truecm] Project Euler (\url{projecteuler.net})
    \end{center}
@@
     $\pi$ to a few thousand digits
+   \end{center}
+  \end{column}%
+ \end{columns}
+
+ \vspace{0.4truecm}
+ {\scriptsize\url{https://github.com/marcinofulus/jupyter4edu/tree/master/augsburg/exercises}}
+\end{frame}
+
+\begin{frame}{Examples of problem sets II}
+
+ \vspace{-0.7truecm}
+ \begin{columns}[t]
+  \begin{column}{0.5\textwidth}
+   \begin{center}
+    \includegraphics[width=0.8\textwidth]{parrondo}
+
+    Parrondo paradoxon
+   \end{center}
+  \end{column}%
+  \begin{column}{0.5\textwidth}
+   \begin{center}
+    \includegraphics[width=0.8\textwidth]{polynomial}
+
+    symbolic manipulation of polynomials
+   \end{center}
+  \end{column}%
+ \end{columns}
+
+ \begin{columns}[t]
+  \begin{column}{0.5\textwidth}
+   \begin{center}
+    \includegraphics[width=0.8\textwidth]{primetime}
+
+    How many primes are times?
+   \end{center}
+  \end{column}%
+  \begin{column}{0.5\textwidth}
+   \begin{center}
+    \includegraphics[width=0.8\textwidth]{birthday}
+
+    birthday problem
    \end{center}

### 2019-09-23 paper/paper.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[smallextended]{svjour3}       % onecolumn (second format)
\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
\usepackage{url}
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
\journalname{Journal of Science Education and Technology}
%
\begin{document}

\title{Experiences in teaching with Jupyter with nbgrader}
%about the article that should go on the front page should be
%placed here. General acknowledgments should be placed at the end of the article.}
% \subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Marcin Kostur \and
        Marina Appiou-Nikiforou \and
        Andreas Papadopoulos \and
	Gert-Ludwig Ingold
}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Marcin Kostur \at
              first address \\
              Tel.: +123-45-678910\\
              Fax: +123-45-678910\\
              \email{marcin.kostur@us.edu.pl}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Marina Appiou-Nikiforou \at
              European University Cyprus \\
              \email{M.Nikiforou@euc.ac.cy}
	   \and
	   Andreas Papadopoulos \at
              European University Cyprus \\
              \email{M.Nikiforou@euc.ac.cy}
	   \and
	   Gert-Ludwig Ingold \at
	      Institut f{\"u}r Physik, Universit{\"a}t Augsburg,
	      Universit{\"a}tsstra{\ss}e 1, 86135 Augsburg, Germany\\
	      \email{gert.ingold@physik.uni-augsburg.de}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
Insert your abstract here. Include keywords, PACS and mathematical
subject classification numbers as needed.
\keywords{First keyword \and Second keyword \and More}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}

Your text comes here. Separate text sections with

\section{Section title}
\label{sec:1}
We discuss Jupyter \cite{jupyter-edu-book} and nbgrader \cite{nbgrader}.

\subsection{Subsection title}
\label{sec:2}

%% For one-column wide figures use
%\begin{figure}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:1}       % Give a unique label
%\end{figure}
%%
%% For two-column wide figures use
%\begin{figure*}
%% Use the relevant command to insert your figure file.
%% For example, with the graphicx package use
%  \includegraphics[width=0.75\textwidth]{example.eps}
%% figure caption is below the figure
%\caption{Please write your figure caption here}
%\label{fig:2}       % Give a unique label
%\end{figure*}
%%
%% For tables use
%\begin{table}
%% table caption is above the table
%\caption{Please write your table caption here}
%\label{tab:1}       % Give a unique label
%% For LaTeX tables use
%\begin{tabular}{lll}
%\hline\noalign{\smallskip}
%first & second & third  \\
%\noalign{\smallskip}\hline\noalign{\smallskip}
%number & number & number \\
%number & number & number \\
%\noalign{\smallskip}\hline
%\end{tabular}
%\end{table}

\begin{acknowledgements}
This work was supported through the Erasmus+ programme of the
European Union.
\end{acknowledgements}

\section*{Conflict of interest}
The authors declare that they have no conflict of interest.

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{99}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
\bibitem{jupyter-edu-book}
Barba, L. A., Barker, L. J., Blank, D. S., Brown, J, Downey, A. B., George, T.,
Heagy, L. J., Mandli, K. T., Moore, J. K., Lippert, D., Niemeyer, K. E.,
Watkins, R. R., West, R. H., Wickes, E., Willing, C., \& Zingale M. (2019).
Teaching and Learning with Jupyter,
\url{https://jupyter4edu.github.io/jupyter-edu-book/}.
\bibitem{nbgrader}
% https://doi.org/10.21105/jose.00032 
Project Jupyter, Blank, D., Bourgin, D., Brown, A., Bussonnier, M.,
Frederic, J., Granger, B., Griffiths, T. L.,  Hamrick, J., Kelley, K.,
Pacer, M., Page, L., P{\'e}rez, F., Ragan-Kelley, B., Suchow, J. W.,
\& Willing, C. (2019).
nbgrader: A Tool for Creating and Grading Assignments in the Jupyter Notebook,
\textit{Journal of Open Source Education}, 2(11), 32.
% Format for books
%\bibitem{RefB}
%Author, Book title, page numbers. Publisher, place (year)
% etc
\end{thebibliography}

\end{document}
% end of file template.tex

### 2019-09-23 paper/paper.tex
Update the adresses
material from my Chorzów presentation added

@@
            Marina Appiou-Nikiforou \at
-              European University Cyprus \\
+              European University Cyprus, 
+	      6 Diogenous Street, 1516 Nicosia, Cyprus\\
               \email{M.Nikiforou@euc.ac.cy}
@@
 	   Andreas Papadopoulos \at
-              European University Cyprus \\
-              \email{M.Nikiforou@euc.ac.cy}
+              European University Cyprus, 
+	      6 Diogenous Street, 1516 Nicosia, Cyprus \\
+              \email{apapado89@gmail.com}
+	     % \emph{Present address:}
 	   \and
@@
 
+\section{Jupyter and Jupyterhub}
+
+pros and cons of using the Jupyter notebook
+\begin{itemize}
+ \item notebook allows to guide students through a problem set
+ \item problems related to out-of-order execution
+ \item \textit{but} students might think that programming in Python
+       necessarily implies working with a Jupyter notebook,
+       they should know about IDEs (spyder, \dots) and
+       editors (vim, \dots)
+\end{itemize}
+
+pros and cons of using Jupyterhub
+\begin{itemize}
+ \item easily accessible interface to problem sets
+ \item no need to install Python on local computer
+ \item consistent working environment for all students
+ \item \textit{but} may be beginners should gather experience with Python
+       on their own computer
+\end{itemize}
+
+\section{\texttt{nbgrader}}
+
+\subsection{Functions}
+
+In \texttt{nbgrader}, the solutions should be implemented in terms of a
+function, so that this function can be tested, typically for different
+parameters. As a consequence, the concept of functions needs to be introduced
+early on in a course. However, the use of functions will be close to the
+concept of a mathematical function in the sense that there will be input values
+and output values. In the tests used to grade the answer, one wants to vary the
+input values and the output value is the object which is to be tested against
+an expectation. Therefore, situations like the absence of arguments or return
+values as well as default arguments and keyword arguments representing more
+advanced concepts, are not required to work with \texttt{nbgrader}.
+
+It is good programming practice to logically structure code into functions of a
+not too large length. Students with \texttt{nbgrader} thus get used to this
+practice early on, in particular if functions are used to build up the final
+solution step by step. The drawback however is that this structure is provided
+to them by the teacher. Students are not actively structuring their code by
+themselves and thus do not really exercise this practise.
+
+In Python, the use of docstrings is highly recommended as way to document
+the behavior of a function. For the use with \texttt{nbgrader}, teachers are
+naturally using docstring because they need to specify the purpose of the
+function and the meaning of the input and output values. Students will thus
+soon consider docstring as standard part of a function. However, they do not
+actively practice the habit of annotating a function with a docstring.
+
+The version of a function released to the students by \texttt{nbgrader} by
+default raises a \texttt{NotImplementedError}. This statement ensures the
+correct working of the function during grading even if no solution is provided.
+Students need to remove the statement and replace it by their solution. It was
+found both in the German as well as the Polish group of students that this need
+was somewhat difficult to convey to them. This difficulty likely arises from
+the fact that students have not been taught the meaning of the \texttt{raise}
+statement. They therefore need to be especially alerted to what they need to do
+with this statement and internalize the removal of the \texttt{raise} statement
+as a recipe.
+
+\subsection{Grading with tests}
+
+The main idea for autograding with \texttt{nbgrader} consists in the teacher
+providing one or several tests while the student needs to develop code which
+satisfies theses tests. This strategy is typical for test-driven development,
+a component of agile programming where tests are written before the code.
+The use of \texttt{nbgrader} thus offers an opportunity to discuss the role
+of testing and the concept of test-driven development. However, it certainly
+depends on the programming background of the specific students whether such
+remarks are fruitful. While one should not expect that students get into the
+habit of test-driven development right away, the concept might stick in their
+mind and be activated at a later time.
+
+The experience in both, the German and the Polish groups was that tests in
+\texttt{nbgrader} problem sets should not only pertain to the specific
+functionality of a solution to be developed by the student. It turned out
+that mistakes like not returning a result or returning a result of an incorrect
+type do not only occur when students start working with \texttt{nbgrader} but
+also in later stages. It has therefore proven useful to separate tests into
+two groups. The first groups of tests check for the existence of a result
+and for the correct type. These are prerequisites for the second group of results
+which test the functionality of the proposed solution for correctness.
+
+The first group of tests should mainly be considered as an opportunity to
+provide feedback to the students. They could be alerted to a missing return
+value by the test
+
+\noindent\texttt{assert result is not None, 'Does your function return
+a result?'}
+
+Quite generally, tests in \texttt{nbgrader} should not only be considered as
+tool for grading but also as an opportunity to give helpful feedback to the
+students while working on the problem set. Providing an expressive error
+message will be helpful for many students who are not yet capable to understand
+or correctly interpret the standard traceback of a Python exception.
+
+A problem very frequently encountered with students results from detailed 
+tests with clear error messages. Students were found to rely in their assessment
+of correctness of their solution almost entirely on the results of the tests.
+In other words, they typically do not develop a critical attitude towards their
+code and do not see the need of doing their own tests. One could imagine to
+have several blocks of tests to demonstrate that the passing of a number of
+tests does not necessarily need that the code is correct. Setting up such a
+test suite however relies on a good knowledge of typical mistakes which can
+be exploited.
+
+\subsection{Addressing heterogeneity}
+
+It is not uncommon that in programming courses one has to address an audience
+with a wide range of programming capabilities. Some students have no
+programming experience whatsoever and might even show a code writing block
+while others have an extensive programming experience, often in a different
+programming language. For students with little or no programming experience,
+more complex exercise problems need to be decomposed into a number steps while
+experienced students might tackle the problems directly without any additional
+help. The latter students often are proficient in a language different from
+Python. A typical indicator are loops over indiced instead of objects contained
+in a list. Here, the feedback mechanism of \texttt{nbgrader} comes handy, in
+particular for \texttt{nbgrader} version 0.6+.
+
+At the same time, the solution of the problem should provide a source of
+satisfaction for all students. When teaching programming to science students, a
+way could be to appeal to the respective scientific background. However, such
+problem sets are often experienced as an unnecessary mental burden.
+
+In exercise work with \texttt{nbgrader}, we have experimented with allowing for
+individual learning paths by providing one main Jupyter notebook and typically
+a few subnotebooks. The main notebook provides the problem description and the
+notebook cells where the final solution should be developed. The subnotebooks
+focus on one aspect of the solution so that students can work towards a complete
+solution step by step.
+
+To illustrate our approach by way of an example, we consider a problem set
+where a graphical representation of a Julia set shall be generated. Students
+need to repetitively carry out an interation $z\rightarrow z^2+c$ where $c$ is
+a given complex constant and the iteration starts from complex values $z_0$
+taken on a grid in the complex plane. The Julia set distinguishes between the
+initial conditions for which $z$ converges from those where $z$ remains bounded.
+In practice, this distinction can be achieved by checking whether an appropriately
+chosen bound is crossed within a given maximum number of iterations. A beautiful
+graphical representation can be obtained by converting the number of iterations
+actually needed to cross the bound into a color.
+
+This problem presents a number of difficulties to the students. From an
+algorithmic point of view, the iteration needs to be implemented and stopped
+once either the bound is crossed or the maximum number of iterations has been
+reached. On a more conceptual level, students were often found to encounter
+problems in imagining how a computer can represent colors. A subnotebook
+explaining the concept of a colormap and allowing students to explore the
+translation of a number into a color representation turns out to be helpful.
+Here, the rich output of the Jupyter notebook is used by providing a SVG
+representation of a square filled with the respective color. Care needs to
+be taken that the students are not irritated by the code needed to provide
+the infrastructure for generating the colored square. The opportunity of
+playing with the conversion of a number into a visual representation of a
+color usually turns out to be very helpful. At this point, a background of
+the RGB or HSV representation of colors is not necessarily required but
+can be offered to interested students in a separate subnotebook.
+
+Another conceptual obstacle for beginners is the idea of pixels in an image
+and how to walk along a two-dimensional grid or even a one-dimensional grid
+within a program. This aspect thus deserves a separate subnotebook and proceeding
+in steps from one to two dimensions proved to be useful. The main reason for
+addressing the one-dimensional case first is due to the observation that if
+students are asked to produce a given number of equally spaced points between
+two values, they typically confuse the number of points with the number of
+intervalls between the points. This issue needs to be first clarified in one
+dimension before the two-dimensional case can be addressed.
+
+Structuring a problem into a number of steps with their own notebooks allows to
+offer individual learning paths. In the supervised exercise sessions it turned
+out that the vast majority of students tends to follow the step-by-step
+learning path which led most directly to the solution. In other words, students
+would have needed very strong encouragement to try to develop the solution
+completely on their own. On the other hand, students typically did not go through
+the additional material provided on the RGB and HSV representation.
+
+On a more technical level, it could be observed that the fact that opening a
+new notebook would lead to an extra tab in the browser could become quite
+confusing for the students unless the take care to close tabs not needed
+anymore. In particular, it happened that the students opened notebooks more
+than once and then getting confused in which notebook they were actually
+working.
+
+Using \texttt{nbgrader} in a scenario as described above, an additional
+difficulty arises in the grading. Students working on subnotebooks should earn
+the respective points even if they are able to only solve some of the steps
+successfully. On the other hand, a student who solves the problem without going
+through all the notebooks should not be punished. The \texttt{nbgrader}
+database holds all the points obtained for the various notebooks, from which a
+final number of points can be determined. This is however a functionality not
+directly provided by \texttt{nbgrader}. If points are assigned appropriately to
+the individual tasks, one can determine the number of points earned as the
+minimum of the total sum of points earned and the maximum number of points
+attributed to the problem set.
+
+\section{Conclusion}
+
 \begin{acknowledgements}

### 2019-09-24 paper/paper.tex
some ideas added to the introduction

@@
 
-Your text comes here. Separate text sections with
+In recent years, Python has become increasingly popular as a first programming
+language at universities. Due to its expressive code, it provides a low entry
+barrier for students while at the same time being well suited for serious
+scientific work (for prominent recent examples see \cite{gw150914} and
+\cite{EHT}). In a teaching environment, the availability of a Python
+distribution for all operating systems likely to be used by students proves to
+be very helpful to ensure a consistent working environment for all students.
+
+The development of the Jupyter notebook has opened up exciting new
+possibilities in teaching \cite{jupyter-edu-book}. At the same time,
+the Jupyter notebook has become the de facto standard among data
+scientists. Employing the Jupyter notebook in teaching provides the
+students not only with new way to learn but also with a tool potentially
+useful for their further career.
+
+While it is possible to run the Jupyter notebook with a Python kernel on the
+students' computer, the fact that work with the Jupyter notebook is browser-based
+suggests to provide students with an access to a Python kernel via a Jupyterhub.
+This approach avoids any installation requirements on the students' side besides
+the availablility of a browser.
+
+For the Jupyter notebook, the extension \texttt{nbgrader} was developed
+\cite{nbgrader} which serves as a tool to distribute, collect and grade problem
+sets for exercises. In the most recent version (0.6.0) at the time of this
+writing, the possibility of giving feedback to students has been significantly
+facilitated.
 
@@
 \label{sec:1}
-We discuss Jupyter \cite{jupyter-edu-book} and nbgrader \cite{nbgrader}.
 
@@
 %
+\bibitem{gw150914}
+% https://doi.org/10.1103/PhysRevD.93.122003
+Abbott, B. P. \textit{et al.} (2016).
+GW150914: First results from the search for binary black hole coalescence with Advanced LIGO,
+\textit{Physical Review D}, 93(12), 122003.
+
+\bibitem{EHT}
+% https://doi.org/10.3847/2041-8213/ab0c57
+The Event Horizon Telescope Collaboration (2019).
+First M87 Event Horizon Telescope Results. III. Data Processing and Calibration,
+\textit{The Astrophysical Journal Letters} 875(1), L3.
+
 \bibitem{jupyter-edu-book}
@@
 \url{https://jupyter4edu.github.io/jupyter-edu-book/}.
+
 \bibitem{nbgrader}
@@
 \end{document}
-% end of file template.tex

### 2019-09-24 paper/paper.tex
title corrected
pros and cons of Jupyter and Jupyterhub formulated

@@
 
-\title{Experiences in teaching with Jupyter with nbgrader}
+\title{Experiences in teaching with Jupyter, Jupyterhub, and nbgrader}
 %about the article that should go on the front page should be
@@
 
-\section{Section title}
-\label{sec:1}
-
-\subsection{Subsection title}
-\label{sec:2}
-
 %% For one-column wide figures use
@@
 \section{Jupyter and Jupyterhub}
+\label{sec:jupyter_jupyterhub}
 
-pros and cons of using the Jupyter notebook
-\begin{itemize}
- \item notebook allows to guide students through a problem set
- \item problems related to out-of-order execution
- \item \textit{but} students might think that programming in Python
-       necessarily implies working with a Jupyter notebook,
-       they should know about IDEs (spyder, \dots) and
-       editors (vim, \dots)
-\end{itemize}
+A Jupyter notebook consists of a collection of two types of cells, namely code
+cells and text cells. The latter may contain text written in markdown, possibly
+including formulas as well as figures. This structure facilitates to guide
+students through a problem set. Usually, the code in a notebook is intended to
+be executed in a linear fashion even though this is not imposed by the
+notebook. In fact, the notebook should invite students to experiment which
+implies the opening of new cells and the execution of code written by the
+student. In this process, out-of-order execution naturally occurs and can have
+side effects which tend to confuse students. While the order of execution of
+individual cells can be inferred by the sequence number attributed to a cell
+once it has been executed, in particular beginners are not aware of the
+problems that out-of-order execution can entail.
 
-pros and cons of using Jupyterhub
-\begin{itemize}
- \item easily accessible interface to problem sets
- \item no need to install Python on local computer
- \item consistent working environment for all students
- \item \textit{but} may be beginners should gather experience with Python
-       on their own computer
-\end{itemize}
+Using the Jupyter notebook for exercises is interesting because students can
+easily experiment with code. At the same time one should be aware that students
+might think that this is the standard way of programming in Python. While it is
+indeed possible to develop pieces of a more complex program in the Jupyter 
+notebook and later combine these pieces, this is not the standard approach for
+complex programming tasks. Therefore, students should also be acquainted with
+the IDEs like Spyder distributed with the Anaconda distribution or the use of
+editors like vim or others which can be configures to support programming in
+Python.
+
+Jupyterhub as a centrally provided resource to run a Jupyter notebook from a
+remote computer makes problem sets or more generally teaching material easily
+accessible without the need for a local Python installation. Furthermore, this
+approach ensures a consistent working environment for all students. However,
+it might be an important experience for students that all work which they
+are doing could in fact be done on their own computer instead of using some
+remote service where the underlying machinery is completely hidden.
 

### 2019-09-24 paper/paper.tex

@@
 \label{intro}
+
+Using technology in everyday teaching is something that we, as instructors, 
+should all consider. Using technology for educational purposes is the key to moderate 
+the gap between students of the new era, students who are being brought up and with 
+technology, and us the instructors of the old era, who were brought up with no internet 
+available and are trying to catch up with technology.
 

### 2019-09-24 paper/paper.tex
collaboration intro -

@@
 
-%% For one-column wide figures use
-%\begin{figure}
-%% Use the relevant command to insert your figure file.
-%% For example, with the graphicx package use
-%  \includegraphics{example.eps}
-%% figure caption is below the figure
-%\caption{Please write your figure caption here}
-%\label{fig:1}       % Give a unique label
-%\end{figure}
-%%
-%% For two-column wide figures use
-%\begin{figure*}
-%% Use the relevant command to insert your figure file.
-%% For example, with the graphicx package use
-%  \includegraphics[width=0.75\textwidth]{example.eps}
-%% figure caption is below the figure
-%\caption{Please write your figure caption here}
-%\label{fig:2}       % Give a unique label
-%\end{figure*}
-%%
-%% For tables use
-%\begin{table}
-%% table caption is above the table
-%\caption{Please write your table caption here}
-%\label{tab:1}       % Give a unique label
-%% For LaTeX tables use
-%\begin{tabular}{lll}
-%\hline\noalign{\smallskip}
-%first & second & third  \\
-%\noalign{\smallskip}\hline\noalign{\smallskip}
-%number & number & number \\
-%number & number & number \\
-%\noalign{\smallskip}\hline
-%\end{tabular}
-%\end{table}
+\section{Introduction new technologies in teaching in mutual collaboraion}
+
+Back in 2017 the collaboration between three partners University of
+Augsburg, European University Cyprus and University of Silesia has
+been initiated in the form of Erasmus+ project. The general aim was to
+research on possibility of implementing good practices in using Python
+programming language and Jupyter notebook in teaching in different
+environments. Experience among parters with both Python and Jupyter
+were diversified. It opened a possibility to observe a process of an
+adotion of new educatioal approach at different phases at the same
+time.
+
+The collaboration consisted on three elements: creation of software
+infrastructure, authoring course materials and experimenting with
+elements of Test Driven Development process using nbgrader.
+
+\texttt{Description of each partner specifics}
 
@@
 
+%% For one-column wide figures use
+%\begin{figure}
+%% Use the relevant command to insert your figure file.
+%% For example, with the graphicx package use
+%  \includegraphics{example.eps}
+%% figure caption is below the figure
+%\caption{Please write your figure caption here}
+%\label{fig:1}       % Give a unique label
+%\end{figure}
+%%
+%% For two-column wide figures use
+%\begin{figure*}
+%% Use the relevant command to insert your figure file.
+%% For example, with the graphicx package use
+%  \includegraphics[width=0.75\textwidth]{example.eps}
+%% figure caption is below the figure
+%\caption{Please write your figure caption here}
+%\label{fig:2}       % Give a unique label
+%\end{figure*}
+%%
+%% For tables use
+%\begin{table}
+%% table caption is above the table
+%\caption{Please write your table caption here}
+%\label{tab:1}       % Give a unique label
+%% For LaTeX tables use
+%\begin{tabular}{lll}
+%\hline\noalign{\smallskip}
+%first & second & third  \\
+%\noalign{\smallskip}\hline\noalign{\smallskip}
+%number & number & number \\
+%number & number & number \\
+%\noalign{\smallskip}\hline
+%\end{tabular}
+%\end{table}
+
 % BibTeX users please use one of

### 2019-09-27 paper/paper.tex

@@
 \texttt{Description of each partner specifics}
+
+At the university of Augsburg, programming courses for students of physics and
+materials science are offered since about ten years. During that time, the
+exercises mentioned in \label{sec:nbgrader} have been gradually developed and
+tested in exercise sessions organized in groups of two or three students under
+guidance of teaching personnel. This format was also be used to develop the
+Jupyter notebooks. The close interaction between students and teachers allowed
+to analyze the way students make use of the notebooks and to identify problems
+in the tests used to do the grading and to improve these tests. While the courses
+so far had about 10-20 students, we expect to have of the order of 100 students
+from next year on. Then, the experience collected with \texttt{nbgrader} will
+be very valuable.
+
+In the lectures on programming, Jupyter notebook have succesfully been used for
+several years for teaching. In addition, students were encouraged to use them
+during the lectures on their laptop but were found to rarely make use of this
+opportunity. They mostly preferred to attentively follow the lecture. However,
+Jupyter notebooks were offered to students for download so that they could
+work on them at home. It is not clear, how much use was made of this possibility.
+On the other hand, in the regular physics courses, Jupyter notebooks have been
+rarely used and probably will not be used for an entire course. However, we
+expect to use Jupyter notebooks more often for demonstration purposes where
+appropriate.
 

### 2019-11-01 paper/paper.tex
project number added to acknowledgments
desc of USl

@@
 were diversified. It opened a possibility to observe a process of an
-adotion of new educatioal approach at different phases at the same
+adoption of new educatioal approach at different phases at the same
 time.
@@
 
-\texttt{Description of each partner specifics}
+\subsection{University of Aubgsburg}
 
@@
 
+\subsection{University of Silesia}
+
+At the University of Silesia Python used to be a a first programming
+language for students of Applied Computer Science. There has beem
+implemented a concept of ``mobile computer classroom'' which consists
+of inexpensive and mobile computers which are are available to
+students in the rental office. Initially it consisted of 10-inch
+netbooks, currently the hardware is based on Chromebooks which are
+very easy to maintain. Therefore an important aspect is the
+accessibility of the work environment in a web browser. Jupyter
+notebook fits perfectly in such a scenario. In recent years several
+work scenarios based on Juyter notebook have been developed in such an
+environment.
+
+Here, we will mainly concentrate on the case of course ``Introduction
+to Artifitial Intelligence``, which is attented by c.a. 50 students
+every year. The course used Jupyter notebook as a teaching tool during
+the lectures and tutorials. Students had to submit their work in the
+form of notebooks form. Additionally automating testing software has been 
+used to help both students as well as teachers.
+
+\subsection{University of Cyprus}
+
 \section{Jupyter and Jupyterhub}
@@
 
-A Jupyter notebook consists of a collection of two types of cells, namely code
-cells and text cells. The latter may contain text written in markdown, possibly
-including formulas as well as figures. This structure facilitates to guide
-students through a problem set. Usually, the code in a notebook is intended to
-be executed in a linear fashion even though this is not imposed by the
-notebook. In fact, the notebook should invite students to experiment which
-implies the opening of new cells and the execution of code written by the
-student. In this process, out-of-order execution naturally occurs and can have
-side effects which tend to confuse students. While the order of execution of
-individual cells can be inferred by the sequence number attributed to a cell
-once it has been executed, in particular beginners are not aware of the
-problems that out-of-order execution can entail.
+Generally speaking, a Jupyter notebook is a document mixing prose,
+code, visualization, together with resources: source code, data,
+media.  It consists of a collection of two types of cells, namely code
+cells and text cells. The latter may contain text written in markdown,
+possibly including formulas as well as figures. This structure
+facilitates to guide students through a problem set. Usually, the code
+in a notebook is intended to be executed in a linear fashion even
+though this is not imposed by the notebook. In fact, the notebook
+should invite students to experiment which implies the opening of new
+cells and the execution of code written by the student. In this
+process, out-of-order execution naturally occurs and can have side
+effects which tend to confuse students. While the order of execution
+of individual cells can be inferred by the sequence number attributed
+to a cell once it has been executed, in particular beginners are not
+aware of the problems that out-of-order execution can entail.
 
@@
 \begin{acknowledgements}
-This work was supported through the Erasmus+ programme of the
-European Union.
+This work was supported by the European Union through Erasmus+ project
+2017-1-PL01-KA203-038747.
 \end{acknowledgements}

### 2019-11-01 paper/paper.tex
i moved out of order discussion to preeliminary section on problems

@@
 possibly including formulas as well as figures. This structure
-facilitates to guide students through a problem set. Usually, the code
-in a notebook is intended to be executed in a linear fashion even
-though this is not imposed by the notebook. In fact, the notebook
-should invite students to experiment which implies the opening of new
-cells and the execution of code written by the student. In this
-process, out-of-order execution naturally occurs and can have side
-effects which tend to confuse students. While the order of execution
-of individual cells can be inferred by the sequence number attributed
-to a cell once it has been executed, in particular beginners are not
-aware of the problems that out-of-order execution can entail.
+facilitates to guide students through a problem set.
 
@@
 
+\section{Problems and pitfalls in using Jupyter notebook}
+Usually, the code in a notebook is intended to be executed in a linear fashion even
+though this is not imposed by the notebook. In fact, the notebook
+should invite students to experiment which implies the opening of new
+cells and the execution of code written by the student. In this
+process, out-of-order execution naturally occurs and can have side
+effects which tend to confuse students. While the order of execution
+of individual cells can be inferred by the sequence number attributed
+to a cell once it has been executed, in particular beginners are not
+aware of the problems that out-of-order execution can entail.
+
 \section{Conclusion}

### 2019-11-04 paper/paper.tex
notebook figure addded
added nbtranslate

@@
 media.  It consists of a collection of two types of cells, namely code
-cells and text cells. The latter may contain text written in markdown,
-possibly including formulas as well as figures. This structure
-facilitates to guide students through a problem set.
+cells and text cells (see fig. \ref{fig:notebook}). The latter may
+contain text written in markdown, possibly including formulas as well
+as figures. This structure facilitates to guide students through a
+problem set.
 
-Using the Jupyter notebook for exercises is interesting because students can
-easily experiment with code. At the same time one should be aware that students
-might think that this is the standard way of programming in Python. While it is
-indeed possible to develop pieces of a more complex program in the Jupyter 
-notebook and later combine these pieces, this is not the standard approach for
-complex programming tasks. Therefore, students should also be acquainted with
-the IDEs like Spyder distributed with the Anaconda distribution or the use of
-editors like vim or others which can be configures to support programming in
-Python.
+\begin{figure}
+  \includegraphics[width=0.5\textwidth]{figs/notebook.png}
+\caption{Jupyter notebooks is a executable document which contains
+  rich-text and code (e.g. in Pyton). }
+\label{fig:notebook}
+\end{figure}
 
-Jupyterhub as a centrally provided resource to run a Jupyter notebook from a
+Jupyter notebook has recenely been popular among researchers as it can
+embody the FAIR (Findable, Accessible, Interoperable, Reusable)
+principles for scholarly communication \cite{Randles2017}. The number
+of notebooks which are publically available on github now exceeds 5
+milion.
+
+Jupyterhub a software providing a way to serve Jupyter notebook for
+multiple users. It can be used in a classes of students, all
+
+ a centrally provided resource to run a Jupyter notebook from a
 remote computer makes problem sets or more generally teaching material easily
@@
 \section{Problems and pitfalls in using Jupyter notebook}
-Usually, the code in a notebook is intended to be executed in a linear fashion even
-though this is not imposed by the notebook. In fact, the notebook
-should invite students to experiment which implies the opening of new
-cells and the execution of code written by the student. In this
-process, out-of-order execution naturally occurs and can have side
-effects which tend to confuse students. While the order of execution
-of individual cells can be inferred by the sequence number attributed
-to a cell once it has been executed, in particular beginners are not
-aware of the problems that out-of-order execution can entail.
+Usually, the code in a notebook is intended to be executed in a linear
+fashion even though this is not imposed by the notebook. In fact, the
+notebook should invite students to experiment which implies the
+opening of new cells and the execution of code written by the
+student. In this process, out-of-order execution naturally occurs and
+can have side effects which tend to confuse students. While the order
+of execution of individual cells can be inferred by the sequence
+number attributed to a cell once it has been executed, in particular
+beginners are not aware of the problems that out-of-order execution
+can entail.
+
+Using the Jupyter notebook for exercises is interesting because
+students can easily experiment with code. At the same time one should
+be aware that students might think that this is the standard way of
+programming in Python. While it is indeed possible to develop pieces
+of a more complex program in the Jupyter notebook and later combine
+these pieces, this is not the standard approach for complex
+programming tasks. Therefore, students should also be acquainted with
+the IDEs like Spyder distributed with the Anaconda distribution or the
+use of editors like vim or others which can be configures to support
+programming in Python.
 
@@
 
+\bibitem{Randles2017}
+B. M. {Randles} and I. V. {Pasquetto} and M. S. {Golshan} and C. L. {Borgman}
+2017 ACM/IEEE Joint Conference on Digital Libraries (JCDL),
+2017, doi={10.1109/JCDL.2017.7991618}
+
 \bibitem{EHT}

### 2020-01-05 paper/paper.tex
some manuscript edits ...

@@
 
-At the University of Silesia Python used to be a a first programming
-language for students of Applied Computer Science. There has beem
-implemented a concept of ``mobile computer classroom'' which consists
-of inexpensive and mobile computers which are are available to
-students in the rental office. Initially it consisted of 10-inch
-netbooks, currently the hardware is based on Chromebooks which are
-very easy to maintain. Therefore an important aspect is the
-accessibility of the work environment in a web browser. Jupyter
-notebook fits perfectly in such a scenario. In recent years several
-work scenarios based on Juyter notebook have been developed in such an
+At the University of Silesia, Python is used as a introductory
+programming language for students of Applied Computer Science. There
+has been implemented a concept of ``mobile computer classroom'' which
+consists of inexpensive and mobile thin-client computers which are are
+available to students in the rental office. They are used to access
+compute resources where the required software is available.
+Chromebooks currently play the role of thins clients, which is why all
+services must be available through a web browser.  Jupyter notebook
+fits perfectly in such a scenario. In recent years several work
+scenarios based on Juyter notebook have been developed in such an
 environment.
 
-Here, we will mainly concentrate on the case of course ``Introduction
-to Artifitial Intelligence``, which is attented by c.a. 50 students
-every year. The course used Jupyter notebook as a teaching tool during
-the lectures and tutorials. Students had to submit their work in the
-form of notebooks form. Additionally automating testing software has been 
-used to help both students as well as teachers.
+In the Institute of Physics, the widespread use of the Python-based
+computer algebra system SageMath\cite{sagemath} system available in
+the browser been started for several years. This turned out to be a
+excellent solution. The Sage notebook (\texttt{sagenb}) provided
+ubiquitous access to computations in science education, and a large
+number of teaching materials where produced. This technology was also
+replicated in partner high schools as part of the iCSE4school project
+(details see \cite{icse4school}). Since 2011, c.a. 2000 students and
+over 500 high school pupils have used the system. The Sage notebooks
+were served using a small computer cluster and until today users
+created around 30000 notebook documents which generated in peak time
+over computation per second.
+
+The notebook system \texttt{sagenb} became obsolete and the currently
+SageMath switched to the Jupyter notebook. Additionally, Jupyter
+notebook supports using many other programming languages. Therefore en
+effort was initiated to replace old \texttt{sagenb} notebooks system
+for the sake of Jupyter, and to expand this technology to other
+teaching programmes.
+
+In this paper, we will mainly concentrate on the case of course
+``Introduction to Artificial Intelligence``, which is attended by
+c.a. 50 Applied Computer Science students every year. The course uses
+Jupyter notebook as a teaching tool during both the lectures and
+tutorials. Students had to submit their work in the form of notebooks
+using \texttt{nbgrader} software. Their work used automated testing to
+help both students as well as teachers.
 
@@
 Jupyterhub a software providing a way to serve Jupyter notebook for
-multiple users. It can be used in a classes of students, all
+multiple users. It can be used in a single classes of students, or can
+be deployed centrally on a faculty level. In both cases it provides
+computational resources to run a Jupyter notebook from a remote
+computing center. It can be used to distribute problem sets or more
+generally teaching material easily. and without the need for a local
+Python installation. Furthermore, this approach ensures a consistent
+working environment for all students.
 
- a centrally provided resource to run a Jupyter notebook from a
-remote computer makes problem sets or more generally teaching material easily
-accessible without the need for a local Python installation. Furthermore, this
-approach ensures a consistent working environment for all students. However,
-it might be an important experience for students that all work which they
-are doing could in fact be done on their own computer instead of using some
-remote service where the underlying machinery is completely hidden.
+However, it might be an important experience for students that all
+work which they are doing could in fact be done on their own computer
+instead of using some remote service where the underlying machinery is
+completely hidden.
 
@@
 % etc
+
+\newcommand{\etalchar}[1]{$^{#1}$}
+\bibitem{sagemath}
+\emph{{S}ageMath, the {S}age {M}athematics {S}oftware {S}ystem ({V}ersion
+  8.9.0)}, The Sage Developers, 2019, {\tt https://www.sagemath.org}.
+
+\bibitem{icse4school}
+  Science education with computational perspective - a project funded by grant from: Erasmus+ Programme 
+ \texttt{http://visual.icse.us.edu.pl/methodology}
+
 \end{thebibliography}

### 2020-01-08 paper/paper.tex
paper fixes...

@@
 At the University of Silesia, Python is used as a introductory
-programming language for students of Applied Computer Science. There
-has been implemented a concept of ``mobile computer classroom'' which
-consists of inexpensive and mobile thin-client computers which are are
-available to students in the rental office. They are used to access
-compute resources where the required software is available.
-Chromebooks currently play the role of thins clients, which is why all
-services must be available through a web browser.  Jupyter notebook
-fits perfectly in such a scenario. In recent years several work
-scenarios based on Juyter notebook have been developed in such an
-environment.
+programming language for students of computer science. There has been
+implemented a concept of ``mobile computer classroom'' which consists
+of inexpensive and portable thin-client computers which are available
+to students in the rental office. They are used to access compute
+resources where the required software is available.  Chromebooks
+currently play the role of thins clients, which is why all services
+must be available through a web browser. Jupyter notebook fits
+perfectly in such a scenario. In recent years several scenarios of
+teaching workflow based on Juyter notebook have been tested and
+implemented.
 
@@
 computer algebra system SageMath\cite{sagemath} system available in
-the browser been started for several years. This turned out to be a
-excellent solution. The Sage notebook (\texttt{sagenb}) provided
+the browser has been in used for several years. This turned out to be
+an excellent solution. The Sage notebook (\texttt{sagenb}) provided
 ubiquitous access to computations in science education, and a large
@@
 over 500 high school pupils have used the system. The Sage notebooks
-were served using a small computer cluster and until today users
-created around 30000 notebook documents which generated in peak time
-over computation per second.
+have been served using a small computer cluster and until today users
+have created over 30000 notebook documents.
 
@@
 tutorials. Students had to submit their work in the form of notebooks
-using \texttt{nbgrader} software. Their work used automated testing to
-help both students as well as teachers.
+using \texttt{nbgrader} software. All excercices include automated
+tests to help both students as well as teachers.
 

### 2020-01-08 paper/paper.tex
editing paper.txt...

@@
 %
+
+\hyphenation{Ju-py-ter} 
 \begin{document}
@@
 At the University of Silesia, Python is used as a introductory
-programming language for students of computer science. There has been
-implemented a concept of ``mobile computer classroom'' which consists
-of inexpensive and portable thin-client computers which are available
-to students in the rental office. They are used to access compute
-resources where the required software is available.  Chromebooks
-currently play the role of thins clients, which is why all services
-must be available through a web browser. Jupyter notebook fits
-perfectly in such a scenario. In recent years several scenarios of
-teaching workflow based on Juyter notebook have been tested and
-implemented.
+programming language for undegraduate students of ``Applied Computer
+Science'' programme. There has been implemented a concept of ``mobile
+computer classroom'' which consists of inexpensive and portable
+thin-client computers which are available to students in the rental
+office. They are used to access compute resources where the required
+software is available.  Chromebooks currently play the role of thins
+clients, which is why all services must be available through a web
+browser. Jupyter notebook fits perfectly in such a scenario. In recent
+years several teaching workflows based on Juyter notebook have been
+tested and implemented.
 
-In the Institute of Physics, the widespread use of the Python-based
-computer algebra system SageMath\cite{sagemath} system available in
-the browser has been in used for several years. This turned out to be
-an excellent solution. The Sage notebook (\texttt{sagenb}) provided
-ubiquitous access to computations in science education, and a large
-number of teaching materials where produced. This technology was also
-replicated in partner high schools as part of the iCSE4school project
-(details see \cite{icse4school}). Since 2011, c.a. 2000 students and
-over 500 high school pupils have used the system. The Sage notebooks
-have been served using a small computer cluster and until today users
-have created over 30000 notebook documents.
-
-The notebook system \texttt{sagenb} became obsolete and the currently
-SageMath switched to the Jupyter notebook. Additionally, Jupyter
-notebook supports using many other programming languages. Therefore en
-effort was initiated to replace old \texttt{sagenb} notebooks system
-for the sake of Jupyter, and to expand this technology to other
+At the Institute of Physics, the Jupyter notebook is not the first
+teaching experience with web based software for mathamatical
+explorations. SageMath\cite{sagemath}, the Python-based computer
+algebra system has been served to students and academic teachers for
+several years. The Sage notebook (\texttt{sagenb}) provided ubiquitous
+access to computations in science education, and a large number of
+teaching materials where produced. This technology was also replicated
+in partner high schools as part of the iCSE4school project (details
+see \cite{icse4school}). Since 2011, c.a. 2000 students and over 500
+high school pupils have used the system. The Sage notebooks have been
+served using a small computer cluster and until today users have
+created over 30000 notebook documents.  Over time, the notebook system
+\texttt{sagenb} has become obsolete and the currently SageMath
+software is available as a kernel to the Jupyter notebook. Therefore
+en effort was initiated to replace old \texttt{sagenb} notebooks
+system for the sake of Jupyter, and to expand this technology to other
 teaching programmes.
 
-In this paper, we will mainly concentrate on the case of course
-``Introduction to Artificial Intelligence``, which is attended by
-c.a. 50 Applied Computer Science students every year. The course uses
-Jupyter notebook as a teaching tool during both the lectures and
-tutorials. Students had to submit their work in the form of notebooks
-using \texttt{nbgrader} software. All excercices include automated
-tests to help both students as well as teachers.
+In this paper, we will the course: ``Introduction to Artificial
+Intelligence``, which is attended by c.a. 50 Applied Computer Science
+students every year. The course uses Jupyter notebook as a main
+teaching tool during both the lectures and tutorials. All materials
+are discributed and collected in as \texttt{*.ipynb} file. Moreover,
+students submit their assignments \texttt{nbgrader} software installed
+on publicly available web-address. All excercices include automated
+tests to help both students in self assessment as well as teachers in
+grading.
 
@@
 \caption{Jupyter notebooks is a executable document which contains
-  rich-text and code (e.g. in Pyton). }
+  rich-text and code (e.g. in Python, R). }
 \label{fig:notebook}
@@
 computing center. It can be used to distribute problem sets or more
-generally teaching material easily. and without the need for a local
+generally teaching material easily and without the need for a local
 Python installation. Furthermore, this approach ensures a consistent
@@
 
-\section{\texttt{nbgrader}}
+\section{Creating and Grading assignments with \texttt{nbgrader}}
 
@@
 
+\section{Typical workflow with Jupyter and nbgrader}
+
+\subsection{Subnotebooks }
+
+\subsection{In Lecture hand out  Perpectron-quizz}
+
+Jupyter notebook is used as hand-out material given during the lecture on learning Perceptron
+
+# Layered neural network
+
+\section{Evaluation}
+
 \section{Problems and pitfalls in using Jupyter notebook}
